# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
library(lattice)
# Read in data from file
activity <- read.csv("./activity.csv")
```


## What is mean total number of steps taken per day?

### Histogram of all daily sums, ignoring days with no data
```{r fig.height=4}
# Get list of daily sums
dailySum <- tapply(activity$steps, activity$date, sum)

# Create histogram of daily sums
histogram(dailySum, type="count", nint=5, xlab = "Daily Sum")
```

### Mean of daily sums
```{r}
# The mean of the daily sums
mean(dailySum, na.rm=TRUE)
```

### Median of daily sums
```{r}
# The median of the daily sums
median(dailySum, na.rm=TRUE)
```


## What is the average daily activity pattern?

### Time plot of average number of steps taken per 5 minute interval over all days
```{r fig.height=4}
# Get the interval averages
averageInterval <- aggregate(steps ~ interval, data=activity, FUN=mean, na.rm=TRUE)

# Create time series plot of average intervals
with(averageInterval, xyplot(steps ~ interval, type="a"))
```

### The 5 minute interval with the maximum average number of steps
```{r}
# Interval which has the maximum average number of steps
averageInterval$interval[which.max(averageInterval$steps)]
```

## Imputing missing values

### Number of missing values
```{r}
# Total number of missing values
missing <- is.na(activity$steps)
sum(missing)
```

### Missing value replacement strategy

Filling in of missing values will be done by using the average across all days
of the corresponding time interval as caluatated in the previous section of
this assignment.

```{r fig.height=4}
# Copy to a new data.frame
activity2 <- activity

# Replace the missing values with the interval average
missingIntervals <- as.character(activity$interval[missing])
row.names(averageInterval) <- averageInterval$interval
activity2$steps[missing] <- averageInterval[missingIntervals, "steps"]
```

### Histogram of all daily sums, ignoring days with no data after n/a replacement
```{r}
# Recalculate the daily sums
dailySum2 <- tapply(activity2$steps, activity2$date, sum)

# Create histogram of new daily sums
histogram(dailySum2, type="count", nint=5, xlab = "Daily Sum")
```

### The mean of the daily sums after n/a replacement
```{r}
# The mean of the new daily sums
mean(dailySum2, na.rm=TRUE)
```

### The median of the daily sums after n/a replacement
```{r}
# The median of the new daily sums
median(dailySum2, na.rm=TRUE)
```

### Impact of missing value replacement strategy on data

The replacement of missing values has changed the data in such a way that it is
now more heavily concentrated around the mean.  The mean has not changed, while
the meadian is now equal to the mean.  This is to be expected as the missing
interval data was replaced with it's average value.

## Are there differences in activity patterns between weekdays and weekends?

### Time plots of average number of steps taken per 5 minute interval over weekdays and weekends 
```{r fig.height=6}

# Get a factor of the days of the week to use to divide the activity2 data
activity2$day <- factor(ifelse(as.POSIXlt(activity2$date)$wday %% 6 == 0, "Weekend", "Weekday"))
                        
# Get the average intervals by weekday/weekend
dayAverage <- aggregate(steps ~ interval:day, data=activity2, FUN=mean)

# Plot the weekend data
with(dayAverage, xyplot(steps ~ interval | day, type="a", layout=c(1,2)))
```